---
title: "XML Side Project"
author: "Group B XML Partners"
date: "June 9, 2019"
output: 
  html_document:  
    toc: true  
    toc_float: true  
    toc_depth: 3  
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning=FALSE)
```

# Intro to XML
Lets start to understand about *Extensible Markup Language*, or XML. 

To put in simple terms, XML is used to **describe data**. 

The XML language is useful for creating unique formats of data and electronically sharing structured data through the public internet and corporate networks. 

The basic foundation of XML code is an *element*, defined by *tags*. If you are unsure what tags are, they are essentially lines of code starting with a beginning tag and ending with an ending tag:

<*beginning tag*> **Element** <*/ending tag*>. 

## Benefits
XML is a good way to store data for several reasons. 

1. XML is a very simple medium to store the data. 
2. It can take large chunks of information and consolidate them into an XML document - meaning that it can effectively organize and provide structure to the user in understanding the data.  
3. XML data is self-describing, meaning that the structure of the data is already embedded with the data. When the data arrives, there is no need to pre-build the structure of the data.

If you're interested in learning more, please visit [this website](https://searchmicroservices.techtarget.com/definition/XML-Extensible-Markup-Language)!

## Description of Data
We have chosen a census dataset from the National Center for Health Statistics, that presents the age-adjusted death rates for the 10 leading causes of death in the US beginning in 1999. Age-adjusted death rates are (per 100,000 population) based on the 2000 US standard population. 

This data enables us to identify health problems in different states, and mainly document the health status of the popluation and of important subgroups. Below, you can find some relevant analyses of our data. 

## Package/functions used to read in xml data
We used XML package in R. XML package has bunch of useful functions for data read in and manipulation. 
1. xmlParse() comes from the XML package in R. It is used commonly to parse or analyze each component of an XML file. This helps to generate an R structure representing the XML tree. 
2. The xmlToDataframe() enables us to extract data from a simple xml document.  
3. The nodes option entails a list of XML nodes (components), which are to be processed. The getNodeSet() enables us to access a specific part of our xml file. 

To learn more about XML package, click [here](https://cran.r-project.org/web/packages/XML/XML.pdf).

# Read/Manipulate Data
To begin reading in XML data, follow these steps:

1. Load in the XML package
2. Apply the xmlParse("Path to file") to begin reading in your XML file. 
3. Grab a node/component from the file using xmlToDataframe
  + This creates a dataframe of factors for each node read in. 
4. Be Creative with how you manipulate your data!

```{r data, echo=TRUE}
library(XML)
library(DT)
library(dplyr)
library(methods)
library(tidyverse)

xmlData <- xmlParse("xmldata.xml")

Y <- setNames(xmlToDataFrame(node=getNodeSet(xmlData, "//year")), "Year")
D <- setNames(xmlToDataFrame(node=getNodeSet(xmlData, "//cause_name")), "Cause")
De <- setNames(xmlToDataFrame(node=getNodeSet(xmlData, "//deaths")), "deaths")
S <- setNames(xmlToDataFrame(node=getNodeSet(xmlData, "//state")), "State")
A <- setNames(xmlToDataFrame(node=getNodeSet(xmlData, "//aadr")), "AADR")

xmlDataDf <- cbind(Y, D, De, S, A)
#creating a dataset

options(digits = 10)
#Create numeric variables
xmlDataDf$AADR <- as.numeric(as.character(xmlDataDf$AADR))
xmlDataDf$deaths <- as.numeric(as.character(xmlDataDf$deaths))
xmlDataDf$Year  <- as.numeric(as.character(xmlDataDf$Year))
# Create Character Variables
xmlDataDf$Cause <- as.character(xmlDataDf$Cause)
xmlDataDf$State <- as.character(xmlDataDf$State)

xmlDataDf2 <- xmlDataDf
str(xmlDataDf2)
```

# Summary Analyses

## 2010 Top 10 leading causes of Death in NC
This datatable presents the top 10 causes of Death in NC proportional to the total number of deaths in the United States. This is an important topic to discuss, because although the number of deaths itself is important, a proportion can give us more information about the deaths in the state of NC. 

A larger proportion can tell us which causes are more affecting residents in NC, which can help us make decisions that benefit the majority of residents in NC suffering from these potential causes. Kidney disease has the highest proportion out of the top 10 causes of death in NC, so focusing on this can be a good step in the future in terms of health care. 

```{r}
library(knitr)
library(ggplot2)

#Create Functions to get info from US and NC
ToGroupUSA <- function(df, na.rm=TRUE, ...) {
  xmlDataDf2 %>% group_by(Year) %>% filter(Year == 2010 & State=="United States" & Cause!= "All causes") %>% arrange(deaths) 
}
ToGroupNC <- function(df1, na.rm=TRUE, ...){
  xmlDataDf2 %>% group_by(Year) %>% filter(Year == 2010 & State=="North Carolina" & Cause!= "All causes") %>% arrange(deaths)
}

xmlUSADataDf <- ToGroupUSA(xmlDataDf2)
names(xmlUSADataDf) <- c("Year", "US-Causes", "US-Deaths", "Location:USA", "USA-AADR")
xmlNCDataDf <- ToGroupNC(xmlDataDf2)
names(xmlNCDataDf) <- c("Year", "NC-Causes", "NC-Deaths", "Location:NC", "NC-AADR")

xmlCombn <- cbind(xmlUSADataDf, xmlNCDataDf) 

# Create a new variable 
options(digits = 5)
xmlCombn1 <- xmlCombn %>% mutate(NCtoUSRatio = `NC-Deaths`/`US-Deaths`) %>% select(Year, `US-Causes`,`NC-Deaths`, `US-Deaths`, NCtoUSRatio) %>% arrange(-NCtoUSRatio)

names(xmlCombn1) <- c("Year", "Causes", "NC-Deaths", "US-Deaths", "NCtoUSRatio")
kable(xmlCombn1, caption = "Comparing National and State deaths")

NCtop3 <- xmlCombn1[1:3, ]
g <- ggplot(NCtop3, aes(x=Causes, y=NCtoUSRatio))
g + geom_bar(stat = "identity", fill="green") + labs(title = "NC Top 3 Causes of Death by Proportion")
```

## Contingency Table 
The Idea of this table is centered around how frequent the State of California will have more than 10k deaths within the 3 year period. The number 10k represents our threshold, which indicates if this is a cause that needs to be focused primarily on. 

Heart disease, Cancer, and Chronic Lower Respiratory Disease (CLRD), each have 2 or more instances of having over 10k deaths. From this data, we can see that these might be the top 3 causes of death that would need major focus on. By devoting more time and money to these fields, we can help a majority of the population prevent and cure these causes. 

```{r Csummaries}
library(knitr)
library(dplyr)

xmlCT <- xmlDataDf2 %>% group_by(Year) %>% select(Year, State, Cause, deaths) %>% filter((State=="California" | State=="Texas" | State=="Florida") & (Year==2008 | Year==2010 | Year==2012) & (Cause!="All causes")) %>% arrange(Year)

GetQuant <- function(df2, na.rm=TRUE, ...) {
  xmlCT %>% mutate(level=ifelse(deaths > 10000, "Over 10k deaths", "Under 10k deaths"))
}

xmlCT1 <- GetQuant(xmlCT)

xmlCT2 <- table(xmlCT1$Year, xmlCT1$Cause, xmlCT1$level)
kable(xmlCT2[1, 1:10, 1:2], align='r', caption = "Occurrences of causes in California with more/less than 10k deaths in 2008, 2010, 2012")

```

## Summary Statistics Table
In these analyses, we are focused on analyzing the 3 largest states and the top leading cause of death in those states, which is heart disease. These tables provide us with the measures of the mean, which in this case refers to the average number of deaths in 2008, 2010, and 2012, and more. We can also analyze the minimum and maximum of those values the deaths as well, and see that the rates of change for Texas and Florida are low, compared to California. 

```{r Nsummaries}
library(knitr)
xmlCAdf <- filter(xmlCT, State=="California" & Cause=="Heart disease")

xmlCAqu <- sapply(xmlCAdf[, 4], quantile)
xmlCAmean <- sapply(xmlCAdf[, 4], mean)
xmlCAcombn <- rbind(xmlCAqu, xmlCAmean)
row.names(xmlCAcombn) <- c("Min", "1st Qu.", "Median", "3rd Qu.", "Max", "Mean")
kable(xmlCAcombn, caption="Summary of Heart Disease Deaths in CA through 3 years")

xmlTXdf <- filter(xmlCT, State=="Texas" & Cause=="Heart disease")

xmlTXqu <- sapply(xmlTXdf[, 4], quantile)
xmlTXmean <- sapply(xmlTXdf[, 4], mean)
xmlTXcombn <- rbind(xmlTXqu, xmlTXmean)
row.names(xmlTXcombn) <- c("Min", "1st Qu.", "Median", "3rd Qu.", "Max", "Mean")
kable(xmlTXcombn, caption="Summary of Heart Disease Deaths in TX through 3 Years")

xmlFLdf <- filter(xmlCT, State=="Florida" & Cause=="Heart disease")

xmlFLqu <- sapply(xmlFLdf[, 4], quantile)
xmlFLmean <- sapply(xmlFLdf[, 4], mean)
xmlFLcombn <- rbind(xmlFLqu, xmlFLmean)
row.names(xmlFLcombn) <- c("Min", "1st Qu.", "Median", "3rd Qu.", "Max", "Mean")
kable(xmlFLcombn, caption="Summary of Heart Disease Deaths in FL through 3 Years")

```

## Graphical Representation of Data Analyses
In this plot, our goal is to find patterns and trends in death and suicide over a 3 year period. We decided to choose these causes, since they are likely the two biggest causes that are less known or less treatable in this list. Focusing on something like this, we hope to obtain valuable data. 

By plotting the age adjusted death rate, our trend shows that California's rates for alzheimer's disease and suicide are not showing any signs of improvement or decline. We see, however for Florida and Texas, that in this 3 year period, health services treating Alzheimer's disease and suicide are improving greatly, with a decline in death rate each year. 

On the other hand, the boxplot with scattered points presents our measure of summary statistics. The end tails points presents our min/max values, the center presents our mean, and the start/end of the boxes presents our quartiles. What this plot ultimately represents is how spread out our data points are (which in this case are the states). We can analyze that in 2008, the death rate was the most spread out, perhaps due to external factors, such as the 2008 financial crisis. Over the years, we see that the means and end point values level off and decrease from 2008 to 2010 to 2012. 
 
```{r plots}
library(ggplot2)
library(dplyr)

options(digits = 10)
xmlData3 <- xmlDataDf2 %>% filter((Year==2008| Year==2010 | Year==2012) & (State=="California" | State=="Texas" | State=="Florida") & (Cause=="Suicide" | Cause== "Alzheimer's disease"))

g <- ggplot(xmlData3, aes(x=as.factor(Year), y=AADR))
g + geom_bar(stat="identity", aes(fill= Cause), position = "dodge") + facet_wrap(~ State) + ylab("Age Adjusted Death Rate") + xlab("Year") + ggtitle("Change in Alzheimer's disease and suicide death rates in 3 year period")

b <- ggplot(xmlData3, aes(x=as.factor(Year), y=AADR))
b + geom_boxplot() + geom_jitter(aes(col=State)) + xlab("Year") + ylab("Age Adjusted Death Rate") + ggtitle("Box Plot of AADR in 3 Largest States in 3 year period")

```


# Alternative way to read in data
We can also read in data straight from the internet!
```{r, echo=TRUE}
library(XML)
library(DT)
library(dplyr)
library(methods)
library(tidyverse)
#The xmlParse() comes from the XML package in R. 
#This is used commonly to parse or analyze each component of an XML file. 
#This helps to generate an R structure representing the XML tree. 
xmlData <- xmlParse("http://data.cdc.gov/api/views/bi63-dtpu/rows.xml")
#extract xml data to dtatframe 
year <- setNames(xmlToDataFrame(node=getNodeSet(xmlData, "//year")), "Year")
cause <- setNames(xmlToDataFrame(node=getNodeSet(xmlData, "//cause_name")), "Cause")
deaths <- setNames(xmlToDataFrame(node=getNodeSet(xmlData, "//deaths")), "Deaths")
state <- setNames(xmlToDataFrame(node=getNodeSet(xmlData, "//state")), "State")
aadr <- setNames(xmlToDataFrame(node=getNodeSet(xmlData, "//aadr")), "AADR")
xmlDataDfA <- cbind(year, cause, deaths, state, aadr)  

options(digits = 10)
#Convert AADR, Deaths & Year to numeric variables
xmlDataDfA$AADR <- as.numeric(as.character(xmlDataDfA$AADR))
xmlDataDfA$Deaths <- as.numeric(as.character(xmlDataDfA$Deaths))
str(xmlDataDfA)
```

# Plots Galore!

## Quantitative Data Analysis 
The table provides us a summary of leading causes of deaths in  Georgia between 2011 and 2016. It shows the total number of deaths and average age adjusted death rate for each cause. We can see that from this table, the average death ratio and average of total deaths are highly correlated.

```{r Summary}
library(knitr)
library(ggplot2)
#Create Functions to get info from US and NC
last5yearSummary <- function(x, na.rm=TRUE, ...) {
  totalDeathsByState <- xmlDataDfA %>% group_by(Cause) %>% filter(Year %in% (2011:2016) & State == x & Cause!= "All causes") %>% mutate(totalDeaths = sum(Deaths), avgAADR = mean(AADR)) %>% select(State, Cause, totalDeaths, avgAADR) 
  return(Last5Year <- distinct(totalDeathsByState))
}
GeorgiaData <- last5yearSummary("Georgia")
#Numeric Summary
knitr::kable(GeorgiaData, caption= "Causes of Death in Georgia in Last 5 years (2011-2016)")
```

## Visual Summary - Quantitative
Bar graph helps us in identifying the leading causes of death in Georgia over period of 2011-2016. The top 3 causes of death were CLRD, Unintentional Injuries followed by stroke. 

The boxplot provides a 5-point summary of the data. By plotting the age adjusted death rate, our trend shows that Maine's rates for heart disease and stroke are consistently higher relative to other New England area. We also see that there was a significant drop in average AADR in 2016 as compared to previous years. 

```{r plot}
library(ggplot2)
library(dplyr)
options(digits = 10)
#Visual Summary
g <- ggplot(GeorgiaData, aes(x=Cause, y=totalDeaths))
g + geom_bar(stat = "identity", aes(fill= Cause)) + labs(title = "Causes of Deaths in Georgia over last 5 years") + scale_color_discrete(name="Cause of Death") + theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank())
SubData <- xmlDataDf %>% filter((Year %in% (2011:2016)) & (State %in% c( "Maine", "Vermont", "New Hampshire", "Massachusetts", "Rhode Island", "Connecticut")) & (Cause %in% c("Stroke" , "Heat disease")))
g1 <- ggplot(SubData, aes(x=as.factor(Year), y=AADR))
g1 + geom_boxplot() + geom_jitter(aes(col=State)) + xlab("Year") + ylab("Age Adjusted Death Rate") + ggtitle("Box Plot of AADR in last 5 years(2011-2016) in New England Area")
```
