---
title: "Project1"
author: "Anchal Saxena"
date: "6/8/2019"
output:
  html_document:
    df_print: paged
    toc: true
    toc_depth: 3
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning=FALSE)
```
# Intro to XML
Describe your type of data (JSON or XML). What is it, where does it get used, and why is it a good
way to store data?
XML stands for *Extensible Markup Language*. 

XML is used to **describe data** and to focus on **what data is**. In XML user can use custom definition, description and structure of data. These datasets then can be stored and shared in the user format across internet and other networks. 

## Benefits
Important characteristics of XML that make it useful in a variety of systems and solutions −

1. XML is extensible − XML allows you to create your own self-descriptive tags, or language, that suits your application.

2. XML carries the data, does not present it − XML allows you to store the data irrespective of how it will be presented.

3. XML is a public standard − XML was developed by an organization called the World Wide Web Consortium (W3C) and is available as an open standard.

For more informstion, visit [this website](https://www.tutorialspoint.com/xml/xml_overview.htm)!

## Description of Data
**NCHS - Leading Causes of Death: United States**, census dataset from the National Center for Health Statistics presents age-adjusted death rates for the 10 leading causes of death in the US beginning in 1999. Age-adjusted death rates are (per 100,000 population) based on the 2000 US standard population.

To get detailed information,click [here](https://catalog.data.gov/dataset/age-adjusted-death-rates-for-the-top-10-leading-causes-of-death-united-states-2013). 

# Read Data
To begin reading in XML data, follow these steps:

1. Load in the XML package
2. Use xmlParse("Path to file") to read in your XML file. 
3. Use xmlToDataframe to create a dataframe for each node in the xml data. 

```{r data, echo=TRUE}
library(XML)
library(DT)
library(dplyr)
library(methods)
library(tidyverse)
#The xmlParse() comes from the XML package in R. 
#This is used commonly to parse or analyze each component of an XML file. 
#This helps to generate an R structure representing the XML tree. 
xmlData <- xmlParse("http://data.cdc.gov/api/views/bi63-dtpu/rows.xml")
#extract xml data to dtatframe 
year <- setNames(xmlToDataFrame(node=getNodeSet(xmlData, "//year")), "Year")
cause <- setNames(xmlToDataFrame(node=getNodeSet(xmlData, "//cause_name")), "Cause")
deaths <- setNames(xmlToDataFrame(node=getNodeSet(xmlData, "//deaths")), "Deaths")
state <- setNames(xmlToDataFrame(node=getNodeSet(xmlData, "//state")), "State")
aadr <- setNames(xmlToDataFrame(node=getNodeSet(xmlData, "//aadr")), "AADR")
xmlDataDf <- cbind(year, cause, deaths, state, aadr)  
str(xmlDataDf)

options(digits = 10)
#Convert AADR, Deaths & Year to numeric variables
xmlDataDf$AADR <- as.numeric(xmlDataDf$AADR)
xmlDataDf$Deaths <- as.numeric(xmlDataDf$Deaths)

str(xmlDataDf)
```

# Data Analysis 

```{r}
library(knitr)
library(ggplot2)
#Create Functions to get info from US and NC
last5yearSummary <- function(x, na.rm=TRUE, ...) {
  totalDeathsByState <- xmlDataDf %>% group_by(Cause) %>% filter(Year %in% (2011:2016) & State == x & Cause!= "All causes") %>% mutate(totalDeaths = sum(Deaths), avgAADR = mean(AADR)) %>% select(State, Cause, totalDeaths, avgAADR)
  return(Last5Year <- distinct(totalDeathsByState))
}

GeorgiaData <- last5yearSummary("Georgia")
#Numeric Summary
knitr::kable(GeorgiaData, caption= "Causes of Death in Georgia in Last 5 years (2011-2016)")
```

# Visual Summary - Quantitave
```{r plots}
library(ggplot2)
library(dplyr)
options(digits = 10)

#Visual Summary
g <- ggplot(GeorgiaData, aes(x=Cause, y=totalDeaths))
g + geom_bar(stat = "identity", aes(fill= Cause)) + labs(title = "Causes of Deaths in Georgia over last 5 years") + scale_color_discrete(name="Cause of Death") + theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank())

SubData <- xmlDataDf %>% filter((Year %in% (2011:2016)) & (State %in% c( "Maine", "Vermont", "New Hampshire", "Massachusetts", "Rhode Island", "Connecticut")) & (Cause %in% c("Stroke" , "Heat disease")))

g1 <- ggplot(SubData, aes(x=as.factor(Year), y=AADR))
g1 + geom_boxplot() + geom_jitter(aes(col=State)) + xlab("Year") + ylab("Age Adjusted Death Rate") + ggtitle("Box Plot of AADR in last 5 years(2011-2016) in New England Area")
```
